Non-parametric approaches to optimal policy are more robust

### 2013-03-06 inst/doc/nonparametric-bayes.Rmd
notes for marc's query

% Carl Boettiger; Marc Mangel; Steve Munch
% Non-parametric approaches to optimal policy are more robust

@@

@@

@@

### (Brief) Discussion of GP inference

* The use of Gaussian processes for inference in dynamical systems [introduced by @Kocijan2005]

@@

_(More thorough, but general-audience targeted.  Technical details and code provided in appendices)._

>>>>

% Non-parametric approaches to optimal policy are more robust

@@

1. Classical approaches are to fix relationship of stock recruitment.  
2. Parameter uncertainty
3. Gaussian processes for model uncertainty

@@

(Big literature here of course, primary approaches have been using
parametric uncertainty and model uncertainty (e.g. belief distribution over models 
I believe. Worth distinguishing our problem from the more trivially 
sized problems in which the action space is discrete and consists
of 2 or three possible choices, e.g. much of the recent Possingham et al work.)

@@

### Background on Gaussian Process inference

The use of Gaussian process regression (known as Kreging in the geospatial literature) to formulate a predictive model is relatively new in the context of modeling dynamical systems [@Kocijan2005] and introduced in the ecological modeling and fisheries management by @Munch2005.

_Is the following  discription useful or too elementary?_

The essense of the Gaussian process approach can be captured in the following thought experiment: An exhaustive parametric approach to the challenge of structural uncertainty might proceed by writing down all possible functional forms for the underlying dynamical system with all possible parameter values for each form, and then consider searching over this huge space to select the most likely model and parameters; or using a Bayesian approach, assign priors to each of these possible models and infer the posterior distribution of possible models. The Gaussian process approach can be thought of as a computationally efficient approximation to this approach. Gaussian processes represent a large class of models that can be though of as capturing or reasonably approximating the set of models in this collection.  By modeling at the level of the process, rather than the level of parametric equation, we can more concisely capture the possible behavior of these curves.  In place of a parametric model of the dynamical system, the Gaussian Process approach postulates a prior distribution of (n-dimensional) curves that can be though of as approximations to a large range of possible

_Do we need more specifics on Gaussian process as an approximation to parametric models? Discussion of Gaussian process vs other machine learning / forecasting approaches that have less clear statistical foundations?  If so, does this all belong in the discussion?_

The Gaussian process is defined by a covariance kernel ... _discussion of our (gaussian) kernel, blah blah how specific here vs appendix?_.  By requiring our kernel to follow a generic functional form, we can compactly describe the Gaussian process using only a handful of parameters (Table 1)

A more thorough introduction to the formulation and use of Gaussian processes can be found in @Rasmussen2006.

parameter       interpretation
---------       -------------- 
$\sigma^2$      The process noise (from the kernel)
$\tau^2$        Variance around the mean
$\beta$_0       The mean is given by a linear model of slope $\beta$ 
$d$             The length-scale of the covariance function
$n$             The observation error

: Table of parameters for the Gaussian process

Rather than estimate values of these parameters directly, we take a hierarchical approach of placing prior distributions on each.  Following @Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise $n$, and exponential prior on the length-scale $d$.

@@

### 2013-03-06 inst/doc/nonparametric-bayes.tex
notes for marc's query

\documentclass[author-year, review]{elsarticle} %review=doublespace preprint=single 5p=2 column
\usepackage{amsmath, amsfonts, amssymb}  % extended mathematics
% My package additions
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
%\linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables

%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in

\usepackage{microtype}
\usepackage{fancyhdr}
\pagestyle{fancy}
\pagenumbering{arabic}

\usepackage{listings}
\lstnewenvironment{code}{\lstset{language=Haskell,basicstyle=\small\ttfamily}}{}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

%%% Syntax Highlighting for code  %%%
%%% Adapted from knitr book %%% 
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{enumerate}
\usepackage{ctable}
\usepackage{float}

% This is needed because raggedright in table elements redefines \\:
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash
\usepackage[normalem]{ulem}
\newcommand{\textsubscr}[1]{\ensuremath{_{\scriptsize\textrm{#1}}}}

% Configure hyperlinks package
\usepackage[breaklinks=true,linktocpage,pdftitle={},pdfauthor={},colorlinks]{hyperref}
\hypersetup{breaklinks=true, pdfborder={0 0 0}}

% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

\VerbatimFootnotes % allows verbatim text in footnotes

% Pandoc header

\begin{document}
\begin{frontmatter}
  \title{}
  \author[cstar]{Carl Boettiger\corref{cor1}}
  \author[noaa]{Stephan B. Munch}
  \author[cstar]{Marc Mangel}
  \ead{cboettig@ucsc.edu}
  \cortext[cor1]{Corresponding author, cboettig@ucsc.edu}
  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 \end{frontmatter}

\section{Non-parametric approaches to optimal policy are more robust}

Carl Boettiger, Marc Mangel, Steve Munch

@@

@@

\subsubsection{(Brief) Discussion of GP inference}

\begin{itemize}
\item
  The use of Gaussian processes for inference in dynamical systems
  {[}introduced by @Kocijan2005{]}
\item
  Gaussian processes in ecological literature (Munch et al. 2005)
\item
  Our methodology (e.g.~following Munch et al. (2005)).
\end{itemize}

@@

\emph{(More thorough, but general-audience targeted. Technical details
and code provided in appendices).}

@@

\subsubsection{Example in a bistable system}

@@

\subsubsection{Discussion of training data}

Both parametric and nonparametric approaches will require some training
data on which to base their model of the process. We generate the
training data under the model described in Eq 1 for 35 time steps, under
a known but not necessarily optimal sequence of harvest intensities,
$h_t$. For simplicity we imagine a fishery that started from zero
harvest pressure and has been gradually increasing the harvest.
(Motivation, alternatives, stationarity, examples without a stable node
(limit-cycle models), examples based on observations near a stable node
alone, and why that isn't impossible).

@@

Which estimates a Ricker model with $r =$

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'alt' not found
\end{verbatim}

, $K =$

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'alt' not found
\end{verbatim}

, and the Allen Allele model with $r =$

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'est' not found
\end{verbatim}

, $K =$

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'est' not found
\end{verbatim}

and $C =$

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'est' not found
\end{verbatim}

.

@@

@@

\bibliography{}

>>>>

\documentclass[author-year, review]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}
\usepackage{lineno} % add 
%\linenumbers % turns line numbering on 
\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%% Redefines the elsarticle footer
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\it \hfill\today}%
 \let\@evenfoot\@oddfoot}
\makeatother

% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\DefineShortVerb[commandchars=\\\{\}]{\|}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Non-parametric approaches to optimal policy are more robust},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

\begin{document}
\begin{frontmatter}
  \title{Non-parametric approaches to optimal policy are more robust}
  \author[cstar]{Carl Boettiger\corref{cor1}}
  \author[noaa]{Stephan B. Munch}
  \author[cstar]{Marc Mangel}
  \ead{cboettig@ucsc.edu}
  \cortext[cor1]{Corresponding author, cboettig@ucsc.edu}
  \address[cstar]{Center for Stock Assessment Research, Department of Applied Math and Statistics, University of California, Mail Stop SOE-2, Santa Cruz, CA 95064, USA}
  \address[noaa]{Southwest Fisheries Science Center, National Oceanic and Atmospheric Administration, 110 Shaffer Road, Santa Cruz, CA 95060, USA}
 \end{frontmatter}

@@

(Big literature here of course, primary approaches have been using
parametric uncertainty and model uncertainty (e.g.~belief distribution
over models I believe. Worth distinguishing our problem from the more
trivially sized problems in which the action space is discrete and
consists of 2 or three possible choices, e.g.~much of the recent
Possingham et al work.)

@@

\subsubsection{Background on Gaussian Process inference}

The use of Gaussian process regression (known as Kreging in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems {[}@Kocijan2005{]} and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005).

\emph{Is the following discription useful or too elementary?}

The essense of the Gaussian process approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters; or
using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
Gaussian process approach can be thought of as a computationally
efficient approximation to this approach. Gaussian processes represent a
large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection. By modeling at the
level of the process, rather than the level of parametric equation, we
can more concisely capture the possible behavior of these curves. In
place of a parametric model of the dynamical system, the Gaussian
Process approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a large range of
possible

\emph{Do we need more specifics on Gaussian process as an approximation
to parametric models? Discussion of Gaussian process vs other machine
learning / forecasting approaches that have less clear statistical
foundations? If so, does this all belong in the discussion?}

The Gaussian process is defined by a covariance kernel \ldots{}
\emph{discussion of our (gaussian) kernel, blah blah how specific here
vs appendix?}. By requiring our kernel to follow a generic functional
form, we can compactly describe the Gaussian process using only a
handful of parameters (Table 1)

A more thorough introduction to the formulation and use of Gaussian
processes can be found in @Rasmussen2006.

\begin{longtable}[c]{@{}ll@{}}
\hline\noalign{\medskip}
parameter & interpretation
\\\noalign{\medskip}
\hline\noalign{\medskip}
$\sigma^2$ & The process noise (from the kernel)
\\\noalign{\medskip}
$\tau^2$ & Variance around the mean
\\\noalign{\medskip}
$\beta$\_0 & The mean is given by a linear model of slope $\beta$
\\\noalign{\medskip}
$d$ & The length-scale of the covariance function
\\\noalign{\medskip}
$n$ & The observation error
\\\noalign{\medskip}
\hline
\noalign{\medskip}
\caption{Table of parameters for the Gaussian process}
\end{longtable}

Rather than estimate values of these parameters directly, we take a
hierarchical approach of placing prior distributions on each. Following
@Gramarcy2005 we use a Gaussian prior on $\beta_0$, an inverse gamma
prior on $\sigma^2$ and $\tau^2$, a gamma prior on the observation noise
$n$, and exponential prior on the length-scale $d$.

\begin{itemize}
\item
  Gaussian processes in ecological literature (Munch et al. 2005)
\item
  Our methodology (e.g.~following Munch et al. (2005)).
\end{itemize}

@@

@@

\subsection{Example in a bistable system}

@@

\subsection{Sample training data}

Both parametric and nonparametric approaches will require training data
on which to base their model of the process. We generate the training
data under the model described in Eq 1 for 35 time steps, under a known
but not necessarily optimal sequence of harvest intensities, $h_t$. For
simplicity we imagine a fishery that started from zero harvest pressure
and has been gradually increasing the harvest.

Using data simulated from a specified model rather than empirical data
permits the comparison against the true underlying dynamics, setting a
bar for the optimal performance possible.

(Motivation, alternatives, stationarity, examples without a stable node
(limit-cycle models), examples based on observations near a stable node
alone, and why that isn't impossible).

@@

Which estimates a Ricker model with $r =$ 1.8501, $K =$ 9.8091, and the
Allen Allele model with $r =$ 2.8079, $K =$ 11.8235 and $C =$ 7.2159.

@@

\section{Acknowledgments}

This work was partially supported by the Center for Stock Assessment
Research, a partnership between the University of California Santa Cruz
and the Fisheries Ecology Division, Southwest Fisheries Science Center,
Santa Cruz, CA and by NSF~grant EF-0924195 to MM.

@@
